Dominance Regions for Rank Two Cluster Algebras

### 2022-05-27 dominance_regions.tex
Boundaries
Partial edits

      \draw (\xmax-1.5,\ymax-1) node{(1.a)};

@@

    \begin{scope}[shift={(13,0)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(1.b)};

@@

    \begin{scope}[shift={(26,0)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(1.c)};

@@

    \begin{scope}[shift={(39, 0)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(1.d)};

@@

    \begin{scope}[shift={(0,-13)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(1.e)};

@@

    \begin{scope}[shift={(13,-13)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(1.f)};

@@

    \begin{scope}[shift={(26,-13)}]

@@

    \begin{scope}[shift={(39,-13)}]

@@

      \draw (\xmax-1.5,\ymax-1) node{(2)};

@@

@@

\begin{remark}
  This result reveals the maximal support possible for a pointed basis which remains pointed under all possible mutations.
  In particular, any candidate basis could be ruled out using these results based upon their support alone.
\end{remark}

The paper is organized as follows.
In Section~\ref{sec:chebyshev}, we collect useful results related to two-parameter Chebyshev polynomials which support our main calculations.
Section~\ref{sec:tropical} contains calculations related to the transformation of $\bfg$-vectors under mutations.
Section~\ref{sec:dominance inequalities} proves Theorem~\ref{th:dominance inequalities}.
Section~\ref{sec:dominance vertices} proves Corollary~\ref{cor:dominance vertices}.
Section~\ref{sec:maximum support} proves Theorem~\ref{th:maximum support}.
The paper ends with Section~\ref{sec:affine} interpreting the dominance polygon in terms of generalized minors in the cases where $b=c=2$.

\section{Chebyshev Polynomials}
  \label{sec:chebyshev}
  Define \emph{two-parameter Chebyshev polynomials} $u_i^\varepsilon$ for $i\in\ZZ$ and $\varepsilon\in\{\pm\}$ recursively by
  \[u_0^\varepsilon=0,\quad u_1^\varepsilon=1,\quad u_{i+1}^\varepsilon=\begin{cases} bu_i^- -u_{i-1}^+ & \text{if $\varepsilon=+$;}\\ cu_i^+-u_{i-1}^- & \text{if $\varepsilon=-$.} \end{cases}\]
    Observe that, by an easy induction, $u_{-i}^\varepsilon=-u_i^\varepsilon$.

  The recursion can be written more succinctly as $u_{i+1}^\varepsilon=u_2^\varepsilon u_i^{-\varepsilon}-u_{i-1}^\varepsilon$ with the additional initial condition $u_2^\varepsilon=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,\ell\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \begin{equation}
      \label{eq:long Chebyshev recursion}
      u_{i+\ell}^\varepsilon=\begin{cases} u_{\ell+1}^\varepsilon u_i^{-\varepsilon}-u_\ell^{-\varepsilon} u_{i-1}^\varepsilon & \text{\upshape if $\ell$ is odd;}\\ u_{\ell+1}^{-\varepsilon} u_i^\varepsilon-u_\ell^\varepsilon u_{i-1}^{-\varepsilon} & \text{\upshape if $\ell$ is even.} \end{cases}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We work by induction on $\ell$, the cases $\ell=0,1$ being tautological and reproducing the defining recursion, respectively.
    Using the claim for $\ell>0$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+\ell+1}^\varepsilon
      &=
      u_{\ell+1}^\varepsilon u_{i+1}^{-\varepsilon}-u_\ell^{-\varepsilon} u_i^\varepsilon
      =(u_2^{-\varepsilon} u_{\ell+1}^\varepsilon-u_\ell^{-\varepsilon}) u_i^\varepsilon-u_{\ell+1}^\varepsilon u_{i-1}^{-\varepsilon}
      =u_{\ell+2}^{-\varepsilon} u_i^\varepsilon - u_{\ell+1}^\varepsilon u_{i-1}^{-\varepsilon}
      \qquad \text{if $\ell$ is odd;}\\ 
      u_{i+\ell+1}^\varepsilon
      &=
      u_{\ell+1}^{-\varepsilon} u_{i+1}^\varepsilon-u_\ell^\varepsilon u_i^{-\varepsilon}
      =(u_2^\varepsilon u_{\ell+1}^{-\varepsilon}-u_\ell^\varepsilon) u_i^{-\varepsilon}-u_{\ell+1}^{-\varepsilon} u_{i-1}^\varepsilon
      =u_{\ell+2}^\varepsilon u_i^{-\varepsilon}-u_{\ell+1}^{-\varepsilon} u_{i-1}^\varepsilon
      \qquad \text{if $\ell$ is even;} 
    \end{align*}
    which gives the claimed recursion for $\ell+1$.
    These calculations can be reversed to show the result for $\ell<0$.
  \end{proof}

>>>>

      \draw (\xmin+1.5,\ymin+1) node{(1.a)};

@@

    \begin{scope}[shift={(14,0)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(1.b)};

@@

    \begin{scope}[shift={(28,0)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(1.c)};

@@

    \begin{scope}[shift={(42, 0)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(1.d)};

@@

    \begin{scope}[shift={(0,-14)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(1.e)};

@@

    \begin{scope}[shift={(14,-14)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(1.f)};

@@

    \begin{scope}[shift={(28,-14)}]

@@

    \begin{scope}[shift={(42,-14)}]

@@

      \draw (\xmin+1.5,\ymin+1) node{(2)};

@@

    %\draw[black,very thick] (-6.5,6.5) -- (45.5,6.5) -- (45.5,-19.5) -- (-6.5,-19.5) -- cycle;
    %\draw[black,very thick] (-6.5,-6.5) -- (45.5,-6.5);
    %\draw[black,very thick] (6.5,6.5) -- (6.5,-19.5);
    %\draw[black,very thick] (19.5,6.5) -- (19.5,-19.5);
    %\draw[black,very thick] (32.5,6.5) -- (32.5,-19.5);
    \draw (21,-22) node{The maximal supports $\cS_\lambda$ as described in Theorem \ref{th:maximum support}.};

@@

The paper is organized as follows.
In Section~\ref{sec:chebyshev}, we collect useful results related to two-parameter Chebyshev polynomials which support our main calculations.
Section~\ref{sec:tropical} contains calculations related to the transformation of $\bfg$-vectors under mutations.
Section~\ref{sec:dominance inequalities} proves Theorem~\ref{th:dominance inequalities}.
Section~\ref{sec:dominance vertices} proves Corollary~\ref{cor:dominance vertices}.
Section~\ref{sec:maximum support} proves Theorem~\ref{th:maximum support}.
The paper ends with Section~\ref{sec:affine} interpreting the dominance polygons in terms of generalized minors in the cases where $b=c=2$.

\section{Chebyshev Polynomials}
  \label{sec:chebyshev}
  Define \emph{two-parameter Chebyshev polynomials} $u_i^\varepsilon$ for $i\in\ZZ$ and $\varepsilon\in\{\pm\}$ recursively by $u_0^\varepsilon=0$, $u_1^\varepsilon=1$, and
  \[u_{i+1}^\varepsilon=\begin{cases} bu_i^- -u_{i-1}^+ & \text{if $\varepsilon=+$;}\\ cu_i^+-u_{i-1}^- & \text{if $\varepsilon=-$.} \end{cases}\]
    Observe that, by an easy induction, $u_{-i}^\varepsilon=-u_i^\varepsilon$.

  %The recursion can be written more succinctly as $u_{i+1}^\varepsilon=u_2^\varepsilon u_i^{-\varepsilon}-u_{i-1}^\varepsilon$ with the additional initial condition $u_2^\varepsilon=\begin{cases} b & \text{if $\varepsilon=+$;}\\ c & \text{if $\varepsilon=-$.} \end{cases}$
  %This leads to the following more general recursive structure.
  \begin{lemma}
    For $i,\ell\in\ZZ$ and $\varepsilon\in\{\pm\}$, we have
    \begin{equation}
      \label{eq:long Chebyshev recursion}
      u_{i+\ell}^\varepsilon=\begin{cases} u_{\ell+1}^\varepsilon u_i^{-\varepsilon}-u_\ell^{-\varepsilon} u_{i-1}^\varepsilon & \text{\upshape if $\ell$ is odd;}\\ u_{\ell+1}^{-\varepsilon} u_i^\varepsilon-u_\ell^\varepsilon u_{i-1}^{-\varepsilon} & \text{\upshape if $\ell$ is even.} \end{cases}
    \end{equation}
  \end{lemma}
  \begin{proof}
    We work by induction on $\ell$, the cases $\ell=0,1$ being tautological and reproducing the defining recursion, respectively.
    Using the claim for $\ell>0$ and then the defining recursion twice, we get
    \begin{align*}
      u_{i+\ell+1}^\varepsilon
      &=
      u_{\ell+1}^\varepsilon u_{i+1}^{-\varepsilon}-u_\ell^{-\varepsilon} u_i^\varepsilon
      =(u_2^{-\varepsilon} u_{\ell+1}^\varepsilon-u_\ell^{-\varepsilon}) u_i^\varepsilon-u_{\ell+1}^\varepsilon u_{i-1}^{-\varepsilon}
      =u_{\ell+2}^{-\varepsilon} u_i^\varepsilon - u_{\ell+1}^\varepsilon u_{i-1}^{-\varepsilon}
      \qquad \text{if $\ell$ is odd;}\\ 
      u_{i+\ell+1}^\varepsilon
      &=
      u_{\ell+1}^{-\varepsilon} u_{i+1}^\varepsilon-u_\ell^\varepsilon u_i^{-\varepsilon}
      =(u_2^\varepsilon u_{\ell+1}^{-\varepsilon}-u_\ell^\varepsilon) u_i^{-\varepsilon}-u_{\ell+1}^{-\varepsilon} u_{i-1}^\varepsilon
      =u_{\ell+2}^\varepsilon u_i^{-\varepsilon}-u_{\ell+1}^{-\varepsilon} u_{i-1}^\varepsilon
      \qquad \text{if $\ell$ is even;} 
    \end{align*}
    which gives the claimed recursion for $\ell+1$.
    These calculations can be reversed to show the result for $\ell<0$.
  \end{proof}
